---
title: "Working RNA-seq data (Pt I)"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: false
    allow_retry: true
runtime: shiny_prerendered
description: "Getting, organizing and annotation of RNA-seq data"
---

## Introduction

```{r setup, include=FALSE}
# SW Carpentry less plans are at https://carpentries-incubator.github.io/bioc-rnaseq/index.html
# Notes for the videos can be found at https://github.com/johnsonra/bioc-rnaseq
library(learnr)
library(RCurl)
library(ai4teaching)

# Set global / environmental variables
if(file.exists(file.path('/var', 'shiny', '.Renviron')))
  readRenviron("/var/shiny/.Renviron")

if(Sys.getenv("ENCRYPTION_KEY") == "")
  Sys.setenv(ENCRYPTION_KEY = OPsecrets::get_secret("ENCRYPTION_KEY", "Private", "learnr", "key_hex"))

Sys.setenv(AI_MODEL = "gemini-3.7-flash")

if(Sys.getenv("GEMINI_API_KEY") == "")
  Sys.setenv(GEMINI_API_KEY = OPsecrets::get_secret("GEMINI_API_KEY", "Private", "Gemini", "ai4teaching_key"))
```

```{r recording-ai-feedback, context="server"}
learnr::event_register_handler("question_submission", function(session, event, data) {
  # data contains:
  # - data$label: Chunk / question ID
  # - data$answer: The raw text/input submitted by the student
  # - data$correct: TRUE/FALSE (from mark_as)
  # - data$message: The custom feedback text returned by mark_as()
  
  # Fallback for message retrieval if nested in data$feedback
  feedback_msg <- if (!is.null(data$message)) {
    data$message
  } else if (!is.null(data$feedback$message)) {
    data$feedback$message
  } else {
    NA_character_
  }

  record <- data.frame(
    student_id  = ifelse(is.null(session$input$student_email), "anonymous", session$input$student_email),
    question_id = data$label,
    submission  = paste(data$answer, collapse = " | "),
    is_correct  = data$correct,
    ai_feedback = feedback_msg,
    timestamp   = as.character(Sys.time()),
    stringsAsFactors = FALSE
  )

  record_answer(record)
})
```

```{r student-info, echo=FALSE}
fluidRow(
  column(6, textInput("student_email", "Email Address:", placeholder = "e.g. me@hood.edu"))
)
```

<!--- 02a --->
<iframe width="560" height="315" src="https://www.youtube.com/embed/Y-f3BAHyK0c?si=enWmX-ypYTVEnfqt" title="Introduction to working with RNA-seq data using Bioconductor" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

<!--
insert a shinychat box here

-   Preamble: I am working through a video tutorial, and I've been asked to complete a task on my own. Please review my response and help me with the task if I need it.
-   Background: We have just been discussing the structure of our project working directory for a bulk RNA-seq analysis, and students have been challenged to create the following directories: `data/`, `scripts/`, `documents/`, and `output/`. It was suggested they work in Positron, as that is what the instructor is using.
-   Student prompt: How did it go? Were you able to create the directory structure?
-->

## Downloading data

<iframe width="560" height="315" src="https://www.youtube.com/embed/AnNetH63y4Y?si=Ouw0ogXCa6sMCNtH" title="Downloading data" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

Add the following files to your download script and add them to your project.

::: {#file-table style="font-size: 65%;"}
| URL | Filename |
|-------------------------------------|-----------------------------------|
| https://github.com/carpentries-incubator/bioc-rnaseq/raw/main/episodes/data/GSE96870_coldata_cerebellum.csv | GSE96870_coldata_cerebellum.csv |
| https://github.com/carpentries-incubator/bioc-rnaseq/raw/main/episodes/data/GSE96870_coldata_all.csv | GSE96870_coldata_all.csv |
| https://github.com/carpentries-incubator/bioc-rnaseq/raw/main/episodes/data/GSE96870_rowranges.tsv | GSE96870_rowranges.tsv |
:::

```{r q1_1, echo=FALSE}
question_text("How big is `GSE96870_rowranges.tsv`?",
              answer_fn(function(value){
                check_answer("How big is the `GSE96870_rowranges.tsv` file?",
                             value,
                             background = "We have just discussed downloading files in R with `download.file`. Students were challenged to download a few data files that we will use in the next section of this module.",
                             solution = "The file should be 3425819 bytes (i.e. 3.3M if they use `ls -lh`). Either answer is sufficient.",
                             model = Sys.getenv("AI_MODEL"),
                             api_key = OPsecrets::get_secret('GEMINI_API_KEY', 'Private', 'Gemini', 'api_key'))
                }), allow_retry = TRUE)
```

<!-- work on this --- pick up here. Decided to do this one in class for Fall 2026, so this is unfinished -->

```{r, echo=FALSE}
chat_ui("my_chat")
```
```{r context="server"}
chat_server("my_chat", callback = my_chat_bot)
```

## RNA-seq quantification & Finding reference sequences

<iframe width="560" height="315" src="https://www.youtube.com/embed/q30FeaV6q04?si=aI3983BuJX9eIRlU" title="RNA-seq quantification" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

Download the latest mouse transcriptome fasta file from [GENCODE](https://www.gencodegenes.org/).

:::{.callout-tip}
-   The filename should end with "transcripts.fa.gz"
-   Consider the `readDNAStringSet()` function from the `Biostrings` package to read the file into R.
:::

```{r q1_1, echo=FALSE}
question_text("What do the entries look like?",
              answer_fn(function(value){
                check_answer("What do the entries look like?",
                             value,
                             background = "We have just downloaded the fasta reference file for the mouse transcriptome. I was asked to load the fasta file using `Biostrings::readDNAStringSet` and explore the data a little - nothing too advanced, though.",
                             solution = "Students may describe the structure of the fasta file (e.g. sequence name followed by the bases in the sequence), or they may describe how it looks when printing the DNAStringSet object to the screen (i.e. shows total number of transcripts - somewhere around 480 thousand total sequences, sequence length for each one, colorful display of the first few and last few bases in the sequenct, etc...). Nothing too specific is required, but they should have looked at the file in some way and have some qualitative comments on it.",
                             model = Sys.getenv("AI_MODEL"),
                             api_key = OPsecrets::get_secret('GEMINI_API_KEY', 'Private', 'Gemini', 'api_key'))
                }), allow_retry = TRUE)
```

## Looking forward

<iframe width="560" height="315" src="https://www.youtube.com/embed/kzO7oN4yVbM?si=jXe-dnJ3HTMbWUWO" title="Looking forward" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>
